---
title: |
  | RCT Summer 2026
  | Assignment 2: Word Count of 10-K Filings over Time
author:
  - name: Fikir Worku Edossa
    email: fikiredossa@googlemail.com
    affiliation: Humboldt-Universität zu Berlin
date: last-modified
date-format: long
execute:
  echo: false
  message: false
  warning: false
abstract: |
  | In this assignment I replicate Panel A of Figure 1 of @DLS2017, which visualizes the increase in 10-K length over time. Using the EDGAR 10-K dataset, I extend the research period of the original study to 2023 and confirm that 10-Ks continue to grow longer. \vspace{6cm}
bibliography: references.bib
format:
  pdf:
    documentclass: article
    number-sections: true
    toc: false
    fig-cap-location: top
fontsize: 11pt
ident: yes
header-includes:
  - \usepackage[nolists]{endfloat}
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \usepackage[hang,flushmargin]{footmisc}
  - \usepackage{caption}
  - \captionsetup[table]{skip=12pt,font=bf}
  - \captionsetup[figure]{skip=12pt,font=bf}
  - \usepackage{array}
  - \usepackage{threeparttable}
  - \usepackage{adjustbox}
  - \usepackage{graphicx}
  - \usepackage{csquotes}
  - \usepackage[margin=1in]{geometry}
  - \usepackage{booktabs}
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(gt)

results       <- readRDS("../output/results.rds")
annual_data   <- results$annual_data
annual_table  <- results$annual_table
length_figure <- results$length_figure
highlights    <- results$highlights
```

\pagebreak

# Introduction

@DLS2017 study the development of 10-K filings over their sample period of 1996 to 2013 and document that the length of the documents increases significantly over time. Using filing metadata from the EDGAR 10-K dataset, I replicate their trend figure and extend the analysis to 2023.

# Procedure

Filing metadata comes from the EDGAR 10-K dataset maintained by the TRR266 server, accessed via DuckDB over HTTPS. I restrict the sample to filings with successful markdown conversion and at least 3,000 words, keep one filing per firm per calendar year (the most recent), and require a filing date between 1995 and 2023. The final sample contains `r format(highlights$sample_size, big.mark = ",")` firm-year observations from `r format(highlights$n_firms, big.mark = ",")` unique firms spanning `r highlights$year_range`.

# Results

```{r}
#| label: fig-length
#| fig-cap: "10-K word count over time. Each point represents the median word count across all firm-year observations in that calendar year."
#| out-width: 85%
length_figure
```

```{r}
#| label: tbl-annual
#| tbl-cap: "Annual summary of 10-K filing characteristics."
annual_table
```

```{r, include=FALSE}
change <- highlights$median_words_last - highlights$median_words_first
pct    <- round(100 * change / highlights$median_words_first)
years  <- strsplit(highlights$year_range, "–")[[1]]
first_year <- years[1]
last_year  <- years[2]
```

@fig-length shows that the time trend documented by @DLS2017 continues beyond their sample period. Median word count rose from `r format(highlights$median_words_first, big.mark = ",")` words in `r first_year` to `r format(highlights$median_words_last, big.mark = ",")` words in `r last_year`, an increase of approximately `r pct` percent.

# Conclusion

Replication using a different data source and extended sample period confirms the main finding of @DLS2017: 10-K filings continue to grow longer over time.

\pagebreak

# References {-}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\noindent
