---
title: "A Minimal Research Project Template"
subtitle: "An Illustrative Example in R"
author: "Course Template"
institute: "Your University"
date: last-modified
date-format: long
fontsize: "10pt"
format:
  beamer:
    slide-level: 3
    number-sections: false
    toc: false
classoption: "aspectratio=169"
execute:
  echo: false
  message: false
  warning: false
header-includes:
  - \usepackage{booktabs}
  - \usepackage{graphicx}
---

```{r setup}
library(knitr)
library(ggplot2)

results <- readRDS("../output/results.rds")
descriptive_table <- results$descriptive_table
scatter_figure    <- results$scatter_figure
table_note        <- results$table_note
highlights        <- results$highlights
```


# Motivation

### About This Template

- A minimal reproducible research workflow in R
- Data flow: **pull** → **prepare** → **analyze** → **present**
- Built on the `mtcars` dataset as a teaching example
- Each stage writes a durable artifact to disk; the paper and presentation read from the same results file


### The Research Question

- How does vehicle weight relate to fuel efficiency?
- Does transmission type matter after controlling for weight?
- Sample: `r highlights$sample_size` vehicles from the `mtcars` dataset
- Average fuel efficiency: `r highlights$avg_mpg` mpg; heaviest vehicle: `r highlights$heaviest_model`


# Data

### Descriptive Statistics

```{r tab_descriptives, results="asis"}
kable(
  descriptive_table,
  format   = "latex",
  booktabs = TRUE,
  linesep  = ""
)
```


# Results

### Weight vs. Fuel Efficiency

```{r fig_scatter, fig.height=4.2}
scatter_figure
```


### Key Findings

- A 1,000 lb increase in weight is associated with a **`r highlights$weight_slope` mpg** change
- Manual transmission cars show a **`r highlights$manual_effect` mpg** difference vs. automatic after controlling for weight
- `r highlights$interpretation`


# Conclusion

### Takeaways

- The template keeps all moving parts visible while staying small enough to understand in one sitting
- Replace the toy `mtcars` data with your own; the workflow shape stays the same
- The paper and this presentation both read from `output/rct-project-template-results.rds` — run the pipeline once, render both outputs
