---
title: "WRDS: Compustat, CRSP, and Linking to 10-K Filings"
subtitle: "Research in Corporate Transparency 2026"
author: "Fikir Worku Edossa"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: false
    code-tools: true
    theme: darkly
    embed-resources: true
    df-print: paged
    mermaid:
      theme: dark
execute:
  echo: true
  eval: true
  warning: false
  message: false
---

<style>
:root {
  color-scheme: dark;
}
body,
.quarto-container,
#quarto-content,
main.content {
  background: #050608 !important;
  color: #e8edf2 !important;
}
a { color: #7cc7ff; }
a:hover { color: #a9dcff; }
.quarto-title-block .subtitle { font-size: 1.25rem; }
.sidebar,
#quarto-sidebar,
.sidebar-navigation,
.sidebar nav[role="doc-toc"] {
  background: #080a0f !important;
  color: #d8dee9 !important;
}
.warning-card {
  border-left: 5px solid #ff6b6b;
  background: #211014;
  color: #f5d8dc;
  padding: 1rem 1.25rem;
  margin: 1rem 0;
}
.info-card {
  border-left: 5px solid #7cc7ff;
  background: #0d1722;
  color: #d6e8f7;
  padding: 1rem 1.25rem;
  margin: 1rem 0;
}
.teacher-note {
  border-left: 5px solid #f5c451;
  background: #1c1708;
  color: #f0e2bd;
  padding: 0.85rem 1.15rem;
  margin: 0.75rem 0 1.5rem 0;
  font-size: 0.92rem;
  border-radius: 0 4px 4px 0;
}
.teacher-note p { margin: 0.35rem 0; }
.teacher-note blockquote {
  border-left: 3px solid #8a7320;
  margin: 0.4rem 0 0.4rem 0.2rem;
  padding: 0.1rem 0 0.1rem 0.8rem;
  color: #ffe9a8;
  font-style: italic;
}
.teacher-note .tn-head {
  text-transform: uppercase;
  letter-spacing: 0.08em;
  font-weight: 700;
  color: #f5c451;
  font-size: 0.78rem;
}
.table-note {
  color: #b8c0cc;
  font-size: 0.9rem;
}
.dataframe { width: 100%; }
.cell-output-display { overflow-x: auto; }
.table,
table,
.dataframe {
  --bs-table-bg: #10131a;
  --bs-table-color: #e8edf2;
  --bs-table-striped-bg: #151a23;
  --bs-table-striped-color: #e8edf2;
  --bs-table-hover-bg: #1b2230;
  color: #e8edf2 !important;
  border-color: #303746 !important;
}
thead,
th {
  background: #182030 !important;
  color: #f2f6fb !important;
}
td {
  color: #e8edf2 !important;
}
pre,
code,
.sourceCode {
  background-color: #0d1117 !important;
  color: #e6edf3 !important;
}
.diagram-block {
  white-space: pre;
  overflow-x: auto;
  background: #0d1117;
  color: #9ad1ff;
  border: 1px solid #30363d;
  border-radius: 4px;
  padding: 1rem;
  font-family: ui-monospace, SFMono-Regular, Menlo, Consolas, "Liberation Mono", monospace;
  font-size: 0.9rem;
  line-height: 1.45;
}
</style>

## What this session covers

By the end of this notebook you will be able to pull two of the most-used datasets in
accounting research from WRDS with code, and join them to the 10-K filings you already met:

- **Compustat** (`comp` library): yearly and quarterly numbers from financial statements (we use the annual data here).
- **CRSP** (`crsp` library): daily and monthly share prices, returns, and volume.

Your team project can lean on four data sources — the prepared **10-K filings**, the
**EDGAR log files**, **Compustat**, and **CRSP**. This session is about the last two, and
about getting all three into one data frame so you can attach financial controls and market reactions to a filing.

## Connecting to WRDS

WRDS hosts Compustat and CRSP on a PostgreSQL database. So we open a local **DuckDB** database and *attach* WRDS to it. DuckDB then becomes our single engine: it can read the remote WRDS tables, read the EDGAR parquet files, and join across them. Load the packages first.

```{r}
#| label: packages
suppressPackageStartupMessages({
  library(DBI)
  library(duckdb)
  library(glue)
  library(dotenv)
  library(httr)
  library(rvest)
  library(dplyr)
  library(ggplot2)
})
```

### Your credentials

Keep it in a `secrets.env` file in the project root with two lines:

```
WRDS_USERNAME=your_wrds_id
WRDS_PASSWORD=your_wrds_password
```

The repo ships `_secrets.env` as an empty template — copy it to `secrets.env` and fill in
your details. `secrets.env` is in `.gitignore`, so git will not commit it. We read it with
`dotenv`.

```{r}
#| label: load-secrets
load_dot_env("../secrets.env")

wrds_user <- Sys.getenv("WRDS_USERNAME")
wrds_pass <- Sys.getenv("WRDS_PASSWORD")

stopifnot(nzchar(wrds_user), nzchar(wrds_pass))
```

### Open DuckDB and attach WRDS

```{r}
#| label: wrds-connect
con <- dbConnect(duckdb::duckdb())

dbExecute(con, "INSTALL postgres; LOAD postgres;")
dbExecute(con, sprintf(paste(
  "SET pg_connection_limit=4;",
  "ATTACH 'dbname=wrds host=wrds-pgdata.wharton.upenn.edu port=9737",
  "user=%s password=%s' AS wrds (TYPE postgres, READ_ONLY)"
), wrds_user, wrds_pass))
```

Check it worked. This just asks DuckDB which databases are attached — it returns instantly without touching WRDS. After attaching, WRDS tables are addressed as `wrds.<library>.<table>`, for example `wrds.comp.company`.

```{r}
#| label: wrds-check
dbGetQuery(con, "SELECT database_name FROM duckdb_databases() WHERE database_name = 'wrds'")
```

<div class="info-card">
To see what's available on WRDS, the command <code>SHOW ALL TABLES</code> lists every
library and table (it's long, so we don't print it here).
</div>

## Compustat and CRSP

We query WRDS by building SQL with `glue_sql()` and running it with `dbGetQuery()`. 

### Compustat: company information (`comp.company`)

`comp.company` has one row per firm, keyed by **GVKEY** (Compustat's permanent firm id). It  also carries each firm's **CIK** — the SEC's id — which we use later to reach the filings.

```{r}
#| label: comp-company
company_q <- glue_sql("
  SELECT gvkey, conm, cik, sic, state
  FROM wrds.comp.company
  WHERE conm LIKE 'A%'
  ORDER BY conm
  LIMIT 5
", .con = con)

dbGetQuery(con, company_q)
```

### Compustat: yearly financials (`comp.funda`)

`comp.funda` holds the annual income statement, balance sheet, and cash-flow items. 

```{r}
#| label: comp-funda
cs_filter <- "consol = 'C' and (indfmt = 'INDL' or indfmt = 'FS')
              and datafmt = 'STD' and popsrc = 'D'"

funda_q <- glue_sql(glue("
  SELECT gvkey, datadate, fyear,
         \"at\", sale, ni, ceq, prcc_f, csho
  FROM wrds.comp.funda
  WHERE {cs_filter}
  LIMIT 5
"), .con = con)

dbGetQuery(con, funda_q)
```

We can take a quick look at the data — for instance, how many firm-years there are per decade.

```{r}
#| label: comp-funda-counts
counts_q <- glue_sql(glue("
  SELECT CAST(floor(fyear / 10) * 10 AS INTEGER) AS decade, COUNT(*) AS n
  FROM wrds.comp.funda
  WHERE {cs_filter}
  GROUP BY decade
  ORDER BY decade
"), .con = con)

dbGetQuery(con, counts_q)
```

### CRSP: prices, returns, and volume

CRSP holds daily and monthly market data for US-listed stocks. Its id is **PERMNO** (one
per security); **PERMCO** is the company-level id. CRSP does **not** use GVKEY — that gap is
what the CCM link table (below) closes. First, a lookup from ticker to PERMNO:

```{r}
#| label: crsp-stocknames
names_q <- glue_sql("
  SELECT permno, comnam, ticker, namedt, nameenddt
  FROM wrds.crsp.stocknames
  WHERE ticker = 'AAPL'
  LIMIT 5
", .con = con)

dbGetQuery(con, names_q)
```

The daily stock file is `crsp.dsf`; the matching market index is `crsp.dsi`, which carries
the value-weighted market return `vwretd` (the benchmark for computing "abnormal" returns).
There are monthly twins, `crsp.msf` and `crsp.msi`, for longer horizons.

```{r}
#| label: crsp-dsf
dsf_q <- glue_sql("
  SELECT permno, date, ret, retx, prc, vol, shrout
  FROM wrds.crsp.dsf
  WHERE permno = 14593 AND date >= DATE '2020-01-01'
  LIMIT 5
", .con = con)

dbGetQuery(con, dsf_q)
```


## Linking the databases

We have three datasets, each with its own identifier:

| Dataset | Its identifier |
|---|---|
| 10-K filings (EDGAR) | CIK |
| Compustat (`comp`) | GVKEY |
| CRSP (`crsp`) | PERMNO |

We need two bridges: GVKEY ↔ PERMNO (to add returns to fundamentals) and CIK ↔ GVKEY (to
attach both to a filing). Because everything lives in one DuckDB connection, we build each
bridge as a **view** and join them in SQL.

### CRSP ↔ Compustat: the CCM link

WRDS ships the **CRSP/Compustat Merged** link in `crsp.ccmxpf_lnkhist`. Each row maps a `gvkey` to a PERMNO (`lpermno`) valid over `[linkdt, linkenddt]`. We keep only clean links — link types `LC`/`LU`/`LS` and primary markers `C`/`P` — and fill a missing `linkenddt` (meaning "still valid") with the latest end date in the table so the date comparison works.

```{r}
#| label: ccm-build
dbExecute(con, "
  CREATE OR REPLACE VIEW ccm_link AS
  SELECT gvkey,
         lpermno AS permno,
         lpermco AS permco,
         linkdt,
         COALESCE(linkenddt, MAX(linkenddt) OVER ()) AS linkenddt
  FROM wrds.crsp.ccmxpf_lnkhist
  WHERE linktype IN ('LC', 'LU', 'LS')
    AND linkprim IN ('C', 'P')
")

dbGetQuery(con, "SELECT * FROM ccm_link LIMIT 5")
```

### EDGAR ↔ Compustat: matching on CIK

A 10-K filing is identified by its **CIK**. The simplest bridge to Compustat is the `cik`
column already in `comp.company`. We cast it to an integer.

```{r}
#| label: cik-gvkey
dbExecute(con, "
  CREATE OR REPLACE VIEW cik_gvkey AS
  SELECT DISTINCT CAST(cik AS INTEGER) AS cik, gvkey
  FROM wrds.comp.company
  WHERE cik IS NOT NULL
")

dbGetQuery(con, "SELECT * FROM cik_gvkey LIMIT 5")
```

<div class="info-card">
<b>Watch the type.</b> Compustat stores <code>cik</code> as zero-padded text
(<code>"0000320193"</code>) while the EDGAR data stores it as a number
(<code>320193</code>). Join text to a number and you silently get <i>zero</i> matches, so we
cast both sides to integer.
</div>

<div class="info-card">
<b>One limitation: </b> <code>comp.company</code> keeps only the firm's
<i>current</i> CIK. A firm that changed CIK over time (often after a restructuring) may miss
its older filings. </div>

### Bringing in the 10-K filings

The 10-K data are the parquet files from `edgar_10k_intro.qmd`, read with DuckDB over HTTPS.
Because they live in the **same** `con` that has WRDS attached, we can join filings to
Compustat and CRSP directly. We expose just the linking columns and cast `cik` to integer.

```{r}
#| label: edgar-view
BASE_URL <- "https://trr266.wiwi.hu-berlin.de/edgar_10k_apr26/"

list_parquet_files <- function(base_url) {
  response <- GET(base_url, timeout(60))
  stop_for_status(response)
  page  <- read_html(content(response, as = "text"))
  links <- html_attr(html_elements(page, "a"), "href")
  links <- links[!is.na(links) & grepl("\\.parquet$", links, ignore.case = TRUE)]
  sort(paste0(base_url, links))
}
duckdb_list_literal <- function(urls) {
  escaped <- gsub("'", "''", urls)
  paste0("['", paste(escaped, collapse = "','"), "']")
}

dbExecute(con, "INSTALL httpfs; LOAD httpfs; SET enable_http_metadata_cache = true;")
file_list <- duckdb_list_literal(list_parquet_files(BASE_URL))

dbExecute(con, sprintf("
  CREATE OR REPLACE VIEW edgar_10k AS
  SELECT TRY_CAST(cik AS INTEGER) AS cik,
         accessionNumber,
         CAST(filingDate AS DATE) AS filingDate,
         form
  FROM parquet_scan(%s, filename = true, union_by_name = true)
", file_list))
```

Now build the linked sample in one SQL join: filings → CIK → GVKEY → PERMNO, requiring the
filing date to fall inside the CCM link window.

```{mermaid}
%%| label: link-diagram
%%| echo: false
flowchart LR
    E["10-K filings<br/>(CIK)<br/><i>text / metadata</i>"]
    C["Compustat<br/>(GVKEY)<br/><i>financials</i>"]
    R["CRSP<br/>(PERMNO)<br/><i>returns</i>"]
    E ---|"cik_gvkey<br/>CIK"| C
    C ---|"ccm_link<br/>GVKEY ↔ PERMNO"| R
    classDef hub fill:#1b8a8f,stroke:#7cc7ff,color:#ffffff;
    class C hub;
```

```{r}
#| label: linked-sample
dbExecute(con, "
  CREATE OR REPLACE TEMP TABLE linked_sample AS
  SELECT e.cik, e.accessionNumber, e.filingDate, g.gvkey, l.permno
  FROM edgar_10k e
  JOIN cik_gvkey g USING (cik)
  JOIN ccm_link  l
    ON l.gvkey = g.gvkey
   AND e.filingDate BETWEEN l.linkdt AND l.linkenddt
  WHERE e.form = '10-K'
    AND e.cik IS NOT NULL
")

dbGetQuery(con, "SELECT * FROM linked_sample LIMIT 5")
```

Report how many filings made it through. Filings that drop out (no current CIK match, foreign issuers, timing gaps) are worth reporting.

```{r}
#| label: match-rate
counts <- dbGetQuery(con, "
  SELECT
    (SELECT COUNT(*) FROM edgar_10k
       WHERE form = '10-K' AND cik IS NOT NULL)         AS n_filings,
    (SELECT COUNT(*) FROM linked_sample)                AS n_with_permno
")

counts |>
  mutate(pct_linked = round(100 * n_with_permno / n_filings, 1))
```

## Reference

The Compustat and CRSP material, and the CRSP–Compustat (CCM) link used here, follow Ian D.
Gow and Tongqing Ding, very helpful textbook [**Empirical Research in Accounting: Tools and Methods**](https://iangow.github.io/far_book/){target="_blank"}
